Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-22 2017-Xudong-SMT/FM-2018-extended.tex
def pLTS

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}

@@

\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThen}[2]{\State \algorithmicif\ #1\ \algorithmicthen\ #2}% \IfThenElse{<if>}{<then>}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}
%\DeclareMathOperator{\card}{card}
%\DeclareMathOperator{\Flat}{Flat}
\renewcommand{\P}{\mathcal P}
\usepackage{listings}
\lstset{basicstyle=\small}
%\usepackage{algorithm} 
%\usepackage{algpseudocode}

@@

\def\Talg{\mathcal{T}_{\Sigma,\P}}
Formally, we assume the existence of a term algebra $\Talg$,
where $\Sigma$ is the signature of the data and action constructors,
and $\P$ a set of variables. Within $\Talg$, we distinguish a set of
data expressions $\mathcal{E}_\P$, including a set of Boolean
expressions $\mathcal{B}_{\P}$ ($\mathcal{B}_{\P}\subseteq\mathcal{E}_\P$).
On top of $\mathcal{E}_\P$ we build the action algebra
$\mathcal{A}_\P$, with $\mathcal{A}_P\subseteq\mathcal{T}_\P,
\mathcal{E}_P\cap\mathcal{A}_P=\emptyset$;
naturally action terms will use data expressions as sub-terms.
The function $\vars(t)$ identifies the set of variables in a term
$t\in\AlgT$.

@@

\begin{definition}
  An \emph{algebra presentation} is a triple $\mathcal{P}=\langle\mathit{Sorts},\mathit{Constrs},\mathit{Ops}\rangle$, where

  \begin{itemize}
  \item $Sorts$ is a set of \emph{sorts}\noteSB{Commented out a footnote here\mdash check if still necessary.}
    %% \footnote{Later we may want to extend this with Sort constructors,
    %%   like Array or Pair, but this is not needed now}
  \item $\mathit{Constrs}$ is a set of \emph{constructor operators}: for each $\mathit{Con} \in \mathit{Constrs}$, $arity(Con)=n \in \mathbb{N}$ is its arity 
    and '$Con : (\mathit{sel}_1,\mathit{sort}_1), \dots, (\mathit{sel}_n,\mathit{sort}_n) \rightarrow \mathit{sort}$' is its signature with the associated selectors.
    For each argument, the pair $(\mathit{sel}_i,\mathit{sort}_i)$ defines an auxiliary
    operator of name $\mathit{sel}_i$ with signature $\mathit{sel}_i : \mathit{sort} \rightarrow \mathit{sort}_i$.
    \item $Ops$ is a set of \emph{auxiliary operators}, with their
      arity and signature, of the form: $Op : \mathit{sort}_1, \dots,  \mathit{sort}_n \rightarrow
      \mathit{sort}$
      \item $\mathit{Constrs}(\mathit{sortname})$ and $\mathit{Sels}(\mathit{sortname})$ are, respectively, the sets of
        constructors and selectors of the sort $\mathit{sortname}$
  \end{itemize}
  \noteSB{I think this can be removed}\hl{All sorts and operator names must be distinct.}
  Among constructors, those of arity 0 are called \emph{constants}:
  $\mathit{Consts}(\mathcal{P}) = \{\mathit{Con} \in \mathit{Constrs}\,|\, \mathit{arity}(\mathit{Con})=0\}$.
\end{definition}

@@

@@

%% \QIN{
%% [The input variables in an action term are those marked with a
%% $\symb{?}$.
%% We additionally suppose that each input variable does not
%% appear somewhere else in the same action term:
%% $p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$]}

\begin{definition}[pLTS]
\label{pLTS}
\noteSB{It's a bit strange that algebra presentations are not referred to here}
A pLTS is a tuple
$pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of \emph{states}.
\item[$\bullet$]
$s_0 \in S$ is the \emph{initial state}.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the \emph{transition relation}, with
$L$ the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterized action, \noteSB{Shouldn't this be $e_b \in \AlgB_{\mathcal{P}}$, for {\em some} $\mathcal{P}$?}\hl{$e_b \in
\AlgB$} is a guard, and the variables $x_j\in P$
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
		$\vars(\alpha)\!\subseteq\! \vars(s)$, 
		\noteSB{I would expect $\vars(e_b)\!\subseteq\! \vars(s)$, since the guard has to be tested on the source state}\hl{$\vars(e_b)\!\subseteq\! \vars(s')$}, and
		$\forall j\!\in\! J .\,\vars(e_j)\!\subseteq\! \vars(s)\land 
x_j\!\in\!\vars(s')$.
%\item[$\bullet$]
The $x_j$ are \emph{state variables} of %% the state
$s'$. 
\end{itemize}
\end{definition}

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure, whereof the leaves are pLTSs and holes:
$\pNet::= pLTS~|~\mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle$,
where
\begin{itemize}
\item[$\bullet$] \noteSB{Why is one $\I$ and the other $\I_\P$?}\hl{$I \in \I$ and $J\in\I_\P$} are the sets over which are indexed sub-pNets and holes, respectively, such that: $I\cap J=\emptyset$ and  $I\cup J\neq\emptyset$.

@@

\item[$\bullet$] $\symb{SV}_k^{k\in K}$, with $K\in\I_\P$, is a set of
  synchronization vectors: for each $k\!\in\! K$,
  $\symb{SV}_k\!=\!\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k$ where
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\! I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$\noteSB{What is the sort of a pNet? Should we define $\Sort(\pNet) \triangleq \{\Label(\symb{SV}_k\,|\,k \in K\}$?}, 
  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!S_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\Label(\symb{SV}_k) = \alpha'_k$.

@@

The semantics of open pNets will be defined as an open automaton, that
is an automaton where each transition composes transitions 
of several \noteSB{These should all be pLTSs, right? Or did I miss something?}\hl{LTSs} with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications. Each state of an open
automaton has a set of \emph{state variables} that can be assigned by
incoming transitions.

@@

\begin{definition}
  Given an algebra presentation $\mathcal{P}_{pnet}=<Sorts,Constrs,Ops>$ and a pNet $pnet$, we
  construct:
  \begin{itemize}
  \item An extended presentation $\EPres_{pnet}<Sorts,Constrs \cup Const(pnet),Ops>$.
  \item For a given open transition $ot$, an environment
    $\EEnv_{pnet,ot}=SVars(pnet) \cup IVars(pnet) \cup FVars(pnet,ot)$. 
  \end{itemize} 
\end{definition}

>>>>

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
%\def\AlgI{\mathcal{I}}

@@

\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThen}[2]{\State \algorithmicif\ #1\ \algorithmicthen\ #2}% \IfThenElse{<if>}{<then>}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}
%\DeclareMathOperator{\card}{card}
%\DeclareMathOperator{\Flat}{Flat}
\usepackage{listings}
\lstset{basicstyle=\small}
%\usepackage{algorithm} 
%\usepackage{algpseudocode}

@@

\renewcommand{\P}{\mathcal V}
\def\Talg{\mathcal{T}_{\Sigma,\P}}
Formally, we assume the existence of a term algebra $\Talg$,
where $\Sigma$ is the signature of the data and action constructors,
and $\P$ a set of variables. Within $\Talg$, we distinguish a set of
data expressions $\mathcal{E}_\P$, including a set of Boolean
expressions $\mathcal{B}_{\P}$ ($\mathcal{B}_{\P}\subseteq\mathcal{E}_\P$).
On top of $\mathcal{E}_\P$ we build the action algebra
$\mathcal{A}_\P$, with $\mathcal{A}_P\subseteq\mathcal{T}_\P,
\mathcal{E}_P\cap\mathcal{A}_P=\emptyset$;
naturally action terms will use data expressions as sub-terms.
The function $\vars(t)$ identifies the set of variables in a term
$t\in\AlgT$.

@@

\def\APres{\mathcal{P}}
\begin{definition}
  An \emph{algebra presentation} is a triple $\APres=\langle\mathit{Sorts},\mathit{Constrs},\mathit{Ops}\rangle$, where

  \begin{itemize}
  \item $Sorts$ is a set of \emph{sorts}\noteSB{Commented out a footnote here\mdash check if still necessary.}
    %% \footnote{Later we may want to extend this with Sort constructors,
    %%   like Array or Pair, but this is not needed now}
  \item $\mathit{Constrs}$ is a set of \emph{constructor operators}: for each $\mathit{Con} \in \mathit{Constrs}$, $arity(Con)=n \in \mathbb{N}$ is its arity 
    and '$Con : (\mathit{sel}_1,\mathit{sort}_1), \dots, (\mathit{sel}_n,\mathit{sort}_n) \rightarrow \mathit{sort}$' is its signature with the associated selectors.
    For each argument, the pair $(\mathit{sel}_i,\mathit{sort}_i)$ defines an auxiliary
    operator of name $\mathit{sel}_i$ with signature $\mathit{sel}_i : \mathit{sort} \rightarrow \mathit{sort}_i$.
    \item $Ops$ is a set of \emph{auxiliary operators}, with their
      arity and signature, of the form: $Op : \mathit{sort}_1, \dots,  \mathit{sort}_n \rightarrow
      \mathit{sort}$
      \item $\mathit{Constrs}(\mathit{sortname})$ and $\mathit{Sels}(\mathit{sortname})$ are, respectively, the sets of
        constructors and selectors of the sort $\mathit{sortname}$
  \end{itemize}
%  \noteSB{I think this can be removed}\hl{All sorts and operator names must be distinct.}
  Among constructors, those of arity 0 are called \emph{constants}:
  $\mathit{Consts}(\APres) = \{\mathit{Con} \in \mathit{Constrs}\,|\, \mathit{arity}(\mathit{Con})=0\}$.
\end{definition}

@@

The definition of an Algebra Presentation, and a set of variables
$\P$ fixes the Term algebra elements $\Talg, \AlgB_{\P}, \AlgA_\P$.

@@

\begin{definition}[pLTS]
\label{pLTS}
%\noteSB{It's a bit strange that algebra presentations are not referred to here}
Given a term algebra $\Talg$, a pLTS is a tuple
$pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of \emph{states}.
\item[$\bullet$]
$s_0 \in S$ is the \emph{initial state}.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the \emph{transition relation}, with
$L$ the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
  where $\alpha \in\AlgA_{\P}$ is a parameterized action,
    $e_b \in \AlgB_{\P}$ is a guard, and the
  variables $x_j\in P$ are assigned the expressions \noteEM{Or rather
    $\AlgE_{\P}\cup\AlgA_{P}$ ?}$e_j\in \AlgE_{\P}$. 
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
%		$\vars(\alpha)\!\subseteq\! \vars(s)$, 
%		\noteSB{I would expect $\vars(e_b)\!\subseteq\!
%                  \vars(s)$, since the guard has to be tested on the
%                 source state}
%                \hl{$\vars(e_b)\!\subseteq\! \vars(s')$}, and
                $\vars(e_b)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\forall j\!\in\! J .\,\vars(e_j)\!\subseteq\! \vars(s)\land 
x_j\!\in\!\vars(s')$.
%\item[$\bullet$]
%The $x_j$ are \emph{state variables} of %% the state $s'$. 
\end{itemize}
\end{definition}

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure, whereof the leaves are pLTSs and holes:
$\pNet::= pLTS~|~\mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle$,
where
\begin{itemize}
\item[$\bullet$] \noteEM{Pire que ca, $\I$ n'est pas defini...}\noteSB{Why is one $\I$ and the other $\I_\P$?}\hl{$I \in \I$ and $J\in\I_\P$} are the sets over which are indexed sub-pNets and holes, respectively, such that: $I\cap J=\emptyset$ and  $I\cup J\neq\emptyset$.

@@

\item[$\bullet$] $\symb{SV}_k^{k\in K}$, with $K\in\I_\P$, is a set of
  synchronization vectors: for each $k\!\in\! K$,
  $\symb{SV}_k\!=\!\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k$ where
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\! I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$
%  \noteSB{What is the sort of a pNet? Should we define $\Sort(\pNet)
%    \triangleq \{\Label(\symb{SV}_k\,|\,k \in K\}$?}
  , 
  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!S_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\Label(\symb{SV}_k) = \alpha'_k$.

@@

The semantics of open pNets will be defined as an open automaton, that
is an automaton where each transition composes transitions 
of several \noteSB{These should all be pLTSs, right? Or did I miss
  something?}\noteEM{Hum, nor really, something intermediate; the actions there can be
  parameterised actions, but the label has no guard and no assignments}\hl{LTSs} with the
actions of some holes; the transition occurs if some predicates hold,
and can involve a set of state modifications. Each state of an open
automaton has a set of \emph{state variables} that can be assigned by
incoming transitions.

@@

\begin{definition}
  Given an algebra presentation $\APres_{pnet}=<Sorts,Constrs,Ops>$ and a pNet $pnet$, we
  construct:
  \begin{itemize}
  \item An extended presentation $\EPres_{pnet}<Sorts,Constrs \cup Const(pnet),Ops>$.
  \item For a given open transition $ot$, an environment
    $\EEnv_{pnet,ot}=SVars(pnet) \cup IVars(pnet) \cup FVars(pnet,ot)$. 
  \end{itemize} 
\end{definition}
